Chord: A Static and Dynamic Program Analysis Framework for Java

### 2010-12-18 main/doc/instr_events.tex
cleaned up and updated documentation

\section{Instrumentation Events}
\label{sec:instr-events}

@@

\begin{enumerate}
\item
Method entry and exit. \\
{\bf Event:} \code{ENTER_METHOD} $m$ $t$ \\
{\bf Desc.:} Generated after thread $t$ enters method $m$ respectively. \\
{\bf Event:} \code{LEAVE_METHOD} $m$ $t$ \\
{\bf Desc.:} Generated before thread $t$ leaves method $m$ respectively.

\item
Loop entry and exit. \\
{\bf Event:} \code{ENTER_LOOP} $w$ $t$ \\
{\bf Desc.:} Generated before thread $t$ enters loop $w$ respectively. \\
{\bf Event:} \code{LEAVE_LOOP} $w$ $t$ \\
{\bf Desc.:} Generated after thread $t$ leaves loop $w$ respectively.

\item
Object allocation \\
{\bf Event:} \code{BEF_NEW} $h$ $t$ \\
{\bf Desc.:} Generated only in the crude trace before thread $t$ executes a \code{new} bytecode instruction at program point $h$. \\
{\bf Event:} \code{AFT_NEW} $h$ $t$ $o$ \\
{\bf Desc.:} Generated only in the crude trace after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
%Note: Both above events are visible only in the crude trace and not in the final trace. They are generated only if argument \code{hasObj} of method {@link #setNewAndNewArrayEvent} is true.
{\bf Event:} \code{NEW} $h$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a \code{new} bytecode instruction at program point $h$ and allocates new object $o$. \\
{\bf Event:} \code{NEW_ARRAY} $h$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a \code{newarray} bytecode instruction at program point $h$ and allocates new object $o$.

\item
Getstatic primitive \\
{\bf Event:} \code{GETSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
{\bf Desc.:} Generated after thread $t$ reads primitive-typed static field $f$ at program point $e$.

\item
Getstatic reference \\
{\bf Event:} \code{GETSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed static field $f$ at program point $e$.

\item
Putstatic primitive \\
{\bf Event:} \code{PUTSTATIC_PRIMITIVE} $e$ $t$ $f$ \\
{\bf Desc.:} Generated after thread $t$ writes primitive-typed static field $f$ at program point $e$.

\item
Putstatic reference \\
{\bf Event:} \code{PUTSTATIC_REFERENCE} $e$ $t$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed static field $f$ at program point $e$.

\item
Getfield primitive \\
{\bf Event:} \code{GETFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
{\bf Desc.:} Generated after thread $t$ reads primitive-typed instance field $f$ of object $b$ at program point $e$.

\item
Getfield reference \\
{\bf Event:} \code{GETFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from reference-typed instance field $f$ of object $b$ at program point $e$.

\item
Putfield primitive \\
{\bf Event:} \code{PUTFIELD_PRIMITIVE} $e$ $t$ $b$ $f$ \\
{\bf Desc.:} Generated after thread $t$ writes primitive-typed instance field $f$ of object $b$ at program point $e$.

\item
Putfield reference \\
{\bf Event:} \code{PUTFIELD_REFERENCE} $e$ $t$ $b$ $f$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to reference-typed instance field $f$ of object $b$ at program point $e$.

\item
Aload primitive \\
{\bf Event:} \code{ALOAD_PRIMITIVE} $e$ $t$ $b$ $i$ \\
{\bf Desc.:} Generated after thread $t$ reads the primitive-typed element at index $i$ of array object $b$ at program point $e$.

\item
Aload reference \\
{\bf Event:} \code{ALOAD_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
{\bf Desc.:} Generated after thread $t$ reads object $o$ from the reference-typed element at index $i$ of array object $b$ at program point $e$.

\item
Astore primitive \\
{\bf Event:} \code{ASTORE_PRIMITIVE} $e$ $t$ $b$ $i$ \\
{\bf Desc.:} Generated after thread $t$ writes the primitive-typed element at index $i$ of array object $b$ at program point $e$.

\item
Astore reference \\
{\bf Event:} \code{ASTORE_REFERENCE} $e$ $t$ $b$ $i$ $o$ \\
{\bf Desc.:} Generated after thread $t$ writes object $o$ to the reference-typed element at index $i$ of array object $b$ at program point $e$.

\item
Method call \\
{\bf Event:} \code{METHOD_CALL} $i$ $t$ \\
{\bf Desc.:} Generated before or after thread $t$ executes the method invocation statement at program point $i$.

\item
Thread start call \\
{\bf Event:} \code{THREAD_START} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{start()} method of \code{java.lang.Thread} at program point $i$ and spawns a thread $o$.

\item
Thread join call \\
{\bf Event:} \code{THREAD_JOIN} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{join()} method of \code{java.lang.Thread} at program point i to join with thread $o$.

\item
Lock acquire \\
{\bf Event:} \code{ACQUIRE_LOCK} $l$ $t$ $o$ \\
{\bf Desc.:} Generated after thread $t$ executes a statement of the form monitorenter $o$ or enters a method synchronized on $o$ at program point $l$.

\item
Lock release \\
{\bf Event:} \code{RELEASE_LOCK} $r$ $t$ $o$ \\
{\bf Desc.:} Generated before thread t executes a statement of the form monitorexit o or leaves a method synchronized on o at program point r.

\item
Thread wait call \\
{\bf Event:} \code{WAIT} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{wait()} method of \code{java.lang.Object} at program point $i$ on object $o$.

\item
Thread notify call \\
{\bf Event:} \code{NOTIFY} $i$ $t$ $o$ \\
{\bf Desc.:} Generated before thread $t$ calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} at program point $i$ on object $o$.
\end{enumerate}

>>>>

\section{Instrumentation events}
\label{sec:instr-events}

@@

\begin{enumerate}
\item
Method entry and exit. \\
{\bf Event:} \code{ENTER_METHOD} m t \\
{\bf Desc.:} Generated after thread t enters method m respectively. \\
{\bf Event:} \code{LEAVE_METHOD} m t \\
{\bf Desc.:} Generated before thread t leaves method m respectively.

\item
Loop entry and exit. \\
{\bf Event:} \code{ENTER_LOOP} w t \\
{\bf Desc.:} Generated before thread t enters loop w respectively. \\
{\bf Event:} \code{LEAVE_LOOP} w t \\
{\bf Desc.:} Generated after thread t leaves loop w respectively.

\item
Object allocation \\
{\bf Event:} \code{BEF_NEW} h t \\
{\bf Desc.:} Generated before thread t executes a \code{new} bytecode instruction at program point h and allocates new object o. \\
{\bf Event:} \code{AFT_NEW} h t o \\
{\bf Desc.:} Generated after thread t executes a \code{new} bytecode instruction at program point h and allocates new object o. \\
{\bf Event:} \code{NEW_ARRAY} h t o \\
{\bf Desc.:} Generated after thread t executes a \code{newarray} bytecode instruction at program point h and allocates new object o.

\item
Getstatic primitive \\
{\bf Event:} \code{GETSTATIC_PRIMITIVE} e t f \\
{\bf Desc.:} Generated after thread t reads primitive-typed static field f at program point e.

\item
Getstatic reference \\
{\bf Event:} \code{GETSTATIC_REFERENCE} e t f o \\
{\bf Desc.:} Generated after thread t reads object o from reference-typed static field f at program point e.

\item
Putstatic primitive \\
{\bf Event:} \code{PUTSTATIC_PRIMITIVE} e t f \\
{\bf Desc.:} Generated after thread t writes primitive-typed static field f at program point e.

\item
Putstatic reference \\
{\bf Event:} \code{PUTSTATIC_REFERENCE} e t f o \\
{\bf Desc.:} Generated after thread t writes object o to reference-typed static field f at program point e.

\item
Getfield primitive \\
{\bf Event:} \code{GETFIELD_PRIMITIVE} e t b f \\
{\bf Desc.:} Generated after thread t reads primitive-typed instance field f of object b at program point e.

\item
Getfield reference \\
{\bf Event:} \code{GETFIELD_REFERENCE} e t b f o \\
{\bf Desc.:} Generated after thread t reads object o from reference-typed instance field f of object b at program point e.

\item
Putfield primitive \\
{\bf Event:} \code{PUTFIELD_PRIMITIVE} e t b f \\
{\bf Desc.:} Generated after thread t writes primitive-typed instance field f of object b at program point e.

\item
Putfield reference \\
{\bf Event:} \code{PUTFIELD_REFERENCE} e t b f o \\
{\bf Desc.:} Generated after thread t writes object o to reference-typed instance field f of object b at program point e.

\item
Aload primitive \\
{\bf Event:} \code{ALOAD_PRIMITIVE} e t b i \\
{\bf Desc.:} Generated after thread t reads the primitive-typed element at index i of array object b at program point e.

\item
Aload reference \\
{\bf Event:} \code{ALOAD_REFERENCE} e t b i o \\
{\bf Desc.:} Generated after thread t reads object o from the reference-typed element at index i of array object b at program point e.

\item
Astore primitive \\
{\bf Event:} \code{ASTORE_PRIMITIVE} e t b i \\
{\bf Desc.:} Generated after thread t writes the primitive-typed element at index i of array object b at program point e.

\item
Astore reference \\
{\bf Event:} \code{ASTORE_REFERENCE} e t b i o \\
{\bf Desc.:} Generated after thread t writes object o to the reference-typed element at index i of array object b at program point e.

\item
Method call \\
{\bf Event:} \code{BEF_METHOD_CALL} i t o \\
{\bf Desc.:} Generated before thread t executes the method invocation statement at program point i with this argument as object o. \\
{\bf Event:} \code{AFT_METHOD_CALL} i t o \\
{\bf Desc.:} Generated after thread t executes the method invocation statement at program point i with this argument as object o. \\
{\bf Note:} These events are not generated before/after constructor calls; use \code{BEF_NEW} and \code{AFT_NEW} events for that purpose.

\item
Thread start call \\
{\bf Event:} \code{THREAD_START} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{start()} method of \code{java.lang.Thread} at program point i and spawns a thread o.

\item
Thread join call \\
{\bf Event:} \code{THREAD_JOIN} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{join()} method of \code{java.lang.Thread} at program point i to join with thread o.

\item
Lock acquire \\
{\bf Event:} \code{ACQUIRE_LOCK} l t o \\
{\bf Desc.:} Generated after thread t executes a statement of the form monitorenter o or enters a method synchronized on o at program point l.

\item
Lock release \\
{\bf Event:} \code{RELEASE_LOCK} r t o \\
{\bf Desc.:} Generated before thread t executes a statement of the form monitorexit o or leaves a method synchronized on o at program point r.

\item
Thread wait call \\
{\bf Event:} \code{WAIT} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{wait()} method of \code{java.lang.Object} at program point i on object o.

\item
Thread notify call \\
{\bf Event:} \code{NOTIFY} i t o \\
{\bf Desc.:} Generated before thread t calls the \code{notify()} or \code{notifyAll()} method of \code{java.lang.Object} at program point i on object o.
\end{enumerate}

### 2010-12-18 main/doc/writing_datalog_analysis.tex
cleaned up and updated documentation

\section{Writing a Datalog Analysis}
\label{sec:writing-datalog-analysis}

@@

\begin{quote}
\begin{verbatim}
# name=datarace-dlog

@@

# Analysis constraints
hasWrite(e1,e2) :- write(e1).
hasWrite(e1,e2) :- write(e2).
candidate(e1,e2) :- field(e1,f), field(e2,f), hasWrite(e1,e2), e1 <= e2.
datarace(t1,e1,t2,e2) :- candidate(e1,e2), reach(t1,e1), reach(t2,e2), \
    alias(e1,e2), escape(e1), escape(e2), unguarded(t1,e1,t2,e2).
\end{verbatim}
\end{quote}

@@

Each Datalog analysis rule is a Horn clause of the form
{\tt R(t) :- R1(t1), ..., Rn(tn)}
meaning that if relations {\tt R1}, ..., {\tt Rn} contain tuples {\tt t1}, ..., {\tt tn}
respectively, then relation {\tt R} contains tuple {\tt t}.
A backslash may be used at the end of a line to break long rules for readability.
The Datalog analysis solver \bddbddb\ used in \Chord\ does not apply any
sophisticated optimizations to simplify the rules; besides the BDD variable order,
the manner in which these rules are expressed heavily affects the performance of
the solver.  For instance, an important manual optimization involves breaking down
long rules into multiple shorter rules communicating via intermediate relations.
See Section \ref{sec:tuning-datalog-analysis} for hints on tuning the performance
of Datalog analyses.

>>>>

\section{Writing a Datalog analysis}
\label{sec:writing-datalog-analysis}

@@

\begin{verbatim}
    # name=datarace-dlog

@@

    # Analysis constraints
    hasWrite(e1,e2) :- write(e1).
    hasWrite(e1,e2) :- write(e2).
    candidate(e1,e2) :- field(e1,f), field(e2,f), hasWrite(e1,e2), e1 <= e2.
    datarace(t1,e1,t2,e2) :- candidate(e1,e2), reach(t1,e1), reach(t2,e2), \
        alias(e1,e2), escape(e1), escape(e2), unguarded(t1,e1,t2,e2).
\end{verbatim}

@@

Each Datalog analysis rule is a Horn clause of the form
{\tt R(t) :- R1(t1), ..., Rn(tn)}
meaning that if relations {\tt R1}, ..., {\tt Rn} contain tuples {\tt t1}, ..., {\tt tn}
respectively, then relation {\tt R} contains tuple {\tt t}.
A backslash may be used at the end of a line to break long rules for readability.
The Datalog analysis solver bddbddb used in Chord does not apply any
sophisticated optimizations to simplify the rules; besides the BDD variable order,
the manner in which these rules are expressed heavily affects the performance of
the solver.  For instance, an important manual optimization involves breaking down
long rules into multiple shorter rules communicating via intermediate relations.
See Section \ref{sec:tuning-datalog-analysis} for hints on tuning the performance
of Datalog analyses.
